\chapter*{Acknowledgments}
I would like to express my gratitude to my supervisors, Luca Frediani, for teaching 
me what I know about quantum chemistry, and Tor Fl\aa, for teaching me the theory of
multiwavelets, and to them both for giving my a lot of freedom to persue my own 
interests and ideas. Luca has had a lot of things going on in his life during these 
years, but his door has always been open (literally, his home front door) if I ever 
needed advice or guidence.

I would like to give my sincere thanks to Jonas Jus\'{e}lius, who I had the privilege
to work with during my master's and the first years of my PhD. I might have known the
basic concepts of programming when I met you, but you turned me into a 
programmer. I miss our sharing of WTFs over a code that doesn't work, and I hope 
we get the chance to work together again soon. 

I would also like to mention the rest of the people at the HPC group who have been most 
accommodating in their support, by tweaking the hardware to suite my needs and by 
helping me cut in line when my jobs were to big for the queue.

I would like to give thanks to the CTCC group in Troms\o, to the people I have shared
office with, Jonas, Arnfinn and Krzysztof (and various random people at our guest spot).
To Peter and Antoine, who is/was also working on the multiwavelet project, and to Marco, 
Arnfinn, Roberto, Maarten, Luca (Oggioni) and Geir who joined me in various classes 
over the years (I have been the only student in several classes in the past, so your 
company has been appreciated), and to the rest of the lunch-eaters in the group. 
Finally, to the people at the CTCC in Oslo for many enjoyable joint meetings.

\pagebreak
\ 
\pagebreak
