\chapter{Summary of papers}
\section{Paper I: Adaptive order polynomial algorithm in a multiwavelet representation scheme}
In this work, a new strategy is presented for the reduction of the storage requirements of
functions in a multiwavelet framework. The work is based on Alpert's\cite{Alpert:1993p5460} definition
of the multiwavelet basis which leads to considerable data compression by allowing adaptive 
refinement of the grid for a given order $k$ of the polynomial basis. We propose an additional 
adaptivity in the polynomial order, where the order $k(n)$ depends on the refinement level $n$.
We have found that decreasing the order with increasing refinement can lead to considerable
reduction in storage requirements for the representations of multivariate functions to a given 
accuracy.  

Stig Rune Jensen wrote the computer implementation of the mathematical formalism presented in the 
paper, and assisted in running the test calculations. The theory was developed by Antoine Durdek.

\pagebreak

\section{Paper II: Linear scaling Coulomb interaction in the multiwavelet basis, a parallel implementation}
The paper describes the implementation of a general Poisson solver in a multiwavelet framework,
using the non-standard form of operators. By exploiting the sparsity in the representation of the
involved functions and operators, we were able to achieve linear scaling complexity with respect 
to system size. The performance of the code was demonstrated for molecular systems with up to
600 atoms. 

The presented code is based on an implementation of the application of operators in the multiwavelet
basis using the non-standard form, written in the C language by Frediani and 
Fossgaard\cite{Frediani:2013p1143}.
The code was completely rewritten in C++ by Stig Rune Jensen and Jonas Jus\`{e}lus using a hybrid 
MPI/OpenMP parallelization strategy. The code, which is called MultiResolution Computational Program 
Package (MRCPP) is organized as a mathematical library with general features such as function 
representation and non-standard operator application in multiple dimensions. Jensen also planned and 
ran all test calculations and wrote parts of the manuscript.

\pagebreak

\section{Paper III: Real-Space Density Functional Theory with Localized Orbitals and Multiwavelets}
We present algorithms for the minimization of the Hartree-Fock and Kohn-Sham energies for many-electron 
molecular systems. The general non-canonical HF/KS equations are rewritten in integral form and solved 
in the multiwavelet framework using localized orbitals. Robust and fast convergence is demonstrated for 
small and medium sized systems, and high accuracy energies are presented for a variety of small molecules.

Stig Rune Jensen wrote (with contributions from Peter Wind) the computational chemistry program
MultiResolution Chemistry (MRChem) based on the MRCPP library, and together with Antoine Durdek,
developed the algorithms for the SCF optimization. Jensen also planned and ran all test calculations 
and wrote parts of the manuscript.


